### 2017-04-25 Writing/draft.tex
Updated outline in draft and remove supplementaries

\subsection{Measuring correlation between characters}

We define characters as being correlated if they give the same phylogenetic information.
In order to measure this similitude in phylogenetic information, we propose a new distance metric to measure the difference between two characters:

@@

\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
$CD$ is a continuous distance metric valid in a mathematical sense (see Appendix 1).
Since we are considering differences as being only Fitch-like (i.e. non-weighted and non-ordinate), we ignore the actual distance between two different character states tokens \citep[i.e. a Gower distance;][]{GowerDist}.
Two same character states tokens have a difference of $0$ and two different ones have a difference of $1$ (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
In order to facilitate and standardise characters states comparison, we standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
In other words, we replaced all the occurrences of the first token to be $0$, the second to be $1$, etc.
This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{0,0,1,2,2,1\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
When the character difference is equal to $0$, it means that characters convey the same phylogenetic signal.
When the character difference is equal to $1$ it means it conveys the most different signal.
For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.

@@

\subsection{Inferring the phylogenies}

@@

\subsection{Comparing the topologies}

@@

\section{Appendix A - Character difference}

The Character Difference ($CD$) metric is a distance $d: CD \times CD \rightarrow [0;\infty)$ with any character $x$, $y$ and $z$ $\in CD$ if an only if the followings are satisfied:

\begin{equation}
    \label{Positivity}
    CD_{(x,y)} \geq 0
\end{equation}
\begin{equation}
    \label{Identity}
    CD_{(x,y)} = 0 \Leftrightarrow x = y
\end{equation}
\begin{equation}
    \label{Equality}
    CD_{x,y} = CD_{y,x}
\end{equation}
\begin{equation}
    \label{Subadditivity}
    CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}
\end{equation}

\subsection{Character difference positivity}

For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ (because the difference between character states is considered as Fitch-like, i.e. the difference between two characters states \texttt{0} and \texttt{8} is equal to $1$ and the difference between two identical character states is equal to $0$):

\begin{equation}
    \label{positive_demo}
    \begin{gathered}
    0 \leq  \abs{x_{n}-y_{n}} \leq 1\\
    0 \leq  \sum_{i=1}^{n}\abs{x_{i}-y_{i}} \leq n\\
    0 \leq  \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} \leq 1\\
    0 \leq  \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} \leq \frac{1}{2}\\
    0 \leq  \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} \leq 1\\
    0 \leq  1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) \leq 1\\
    \end{gathered}
\end{equation}

Thus equation \ref{Positivity} is true.

\subsection{Character difference identity}

For any $n \in [1;\infty)$ and $x_{n} = y_{n}$ then:
%TG: Is that valid for a demonstration?
\begin{equation}
    \label{identitydemo}
    \begin{aligned}
    \abs{x_{n}-y_{n}} &= 0\\
    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= 0\\
    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} &= 0\\
    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \frac{1}{2}\\
    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= 1\\
    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) &= 0\\
    \end{aligned}
\end{equation}

Thus equation \ref{Identity} is true.

\subsection{Character difference equality}

For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$, similarly to equation \ref{positive_demo}:

\begin{equation}
    \label{equality_demo}
    \begin{aligned}
    \abs{x_{n}-y_{n}} &= \abs{y_{n}-x_{n}}\\
    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= \sum_{i}^{n}\abs{y_{i}-x_{i}}\\
    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} &= \frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}\\
    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}\\
    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= \frac{\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\\
    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) &= 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)\\
    \end{aligned}
\end{equation}

Thus equation \ref{Equality} is true.

\subsection{Character difference subadditivity}

For any $n, j, b \in [1;\infty)$ and $x_{n}$, $y_{n}$, $z_{n}$ $\in (0,1)$:

\begin{equation}
    \label{subadditivity_demo}
    \begin{aligned}
    CD_{x,y} &\leq CD_{x,z} + CD_{z,y}\\
    \text{if:}\\
    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
     &\leq 
     1 - \left(\frac{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}}}{\frac{1}{2}}\right) 
     +
     1 - \left(\frac{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}}}{\frac{1}{2}}\right) \\
     \text{then, the opposite is:}\\
    \underbrace{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} }_{\in[0,1/2]}
     &\geq 
     \underbrace{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}} }_{\in[0,1/2]}
     +
     \underbrace{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}} }_{\in[0,1/2]} - 2\\
     \text{thus:}\\
     \in[0,\frac{1}{2}] &\geq \in[0,\frac{1}{2}] \in[0,\frac{1}{2}] (-2) \\
     \in[0,\frac{1}{2}] &\geq \in[-2,-1]\\
    \end{aligned}
\end{equation}

Thus equation \ref{Subadditivity} is true.

>>>>

We applied the following simulation protocol to assess the effect of characters correlation (Fig.\ref{Fig:outline}):
\begin{enumerate}
    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices.
    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by duplicating characters in order to maximise and minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters. We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximised or minimised character difference to create null expectancy matrices (hereafter called the ``null'' matrices).
    \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``null'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``null'' trees).
    \item \textbf{Comparing topologies}: finally, we compared to topologies in two ways: (1) first we compared the ``normal'' to the ``null'' trees to assess our simulation protocol; (2) second, we compared the ``maximised'' and ``minimised'' trees to the ``null'' and ``normal'' trees respectively to measure the signal to noise ratio and thus measure the effect of character correlation on topology.
\end{enumerate}
Each steps are described below in more details, as well as our proposed definition of the Character Difference metric.

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using Maximum Parsimony and Bayesian inference (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).}
\label{Fig:outline}
\end{figure}

\subsection{Measuring correlation between characters}
We define characters as being correlated if they give the same phylogenetic information.
In order to measure this similitude in phylogenetic information, we propose a new distance metric to measure the difference between two characters:

@@

\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
$CD$ is a continuous distance metric valid in a mathematical sense (see Supplementary).
Since we are considering differences as being only Fitch-like (i.e. non-weighted and non-ordinate), we ignore the actual distance between two different character states tokens \citep[i.e. a Gower distance;][]{GowerDist}.
Two same character states tokens have a difference of $0$ and two different ones have a difference of $1$ (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
In order to facilitate and standardise characters states comparison, we standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
In other words, we replaced all the occurrences of the first token to be $0$, the second to be $1$, etc.
This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{0,0,1,2,2,1\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
When the character difference is equal to $0$, it means that characters convey the same phylogenetic signal.
When the character difference is equal to $1$ it means it conveys the most different signal.
For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.

@@

\subsection{Inferring topologies}

@@

\subsection{Comparing topologies}

@@
